---
title: "SMPA - Grouping Assignment"
author: "Nikola Misetic"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
---

```{r}

#load libraries
library(tidyverse)
library(lubridate)

#load data
payments <- readRDS("data/trump_campaign_payments.rds")

```

## Introduction

**What data will you be working with?**

A dataset of disbursements (payments and purchases) by the Donald J. Trump for President campaign committee during the 2020 election cycle. The included payments are from 2017 and 2018.

This data and examples were developed as part of the NYT Open training, which we've adapted from here. Note that the dataset has been alterted for instructional purposes, so doesn't represent the original FEC data in its entirety.

In the environment tab in the upper right, click on the `payments` dataset to open the RStudio viewer to examine the data. You can also run the chunk below as well to bring up a slice of the data.

```{r}

head(payments)


```

## Before you begin

Take a moment to look at the various categories included for each disbursement.

-   Notice that there are several date-related fields:

    -   FEC election cycle

    -   report year

    -   election type

    -   disbursement date

-   There are a few fields about the recipient. Along with city and state, you have:

    -   the recipient name

    -   the payee's first and last name (an individual who was paid for the "recipient") and

    -   the entity type (individual v. organization)

## Initial Narrative Questions

Without doing any analysis yet, what are three questions you might try to answer using this dataset? Write your narrative answers in the spaces below.

1.  Which groups did the Trump campaign give most of their money too?

2.  What is the correlation between specific election cycle and amount of money given?

3.  Which cities received the most money?

## Coding Assignment Questions

**Please write your R tidyverse code to help answer the questions below in the provided R chunks. You can make more than one chunk if you prefer.**

**Also, describe what you're doing either in text out "in the open" above the chunk, or in comments (prefaced by #) in the R chunk itself.**

**You'll need to use grouping, filtering, or a combination of both to find the answers. Remember that R is case-sensitive when you refer to any text values within the columns**

<br>

How much was spent in the 2017 vs. 2018 reporting year?

```{r}

payments %>%
  group_by(report_year) %>% 
  summarize(tot_money = sum(disbursement_amount))
  
# I grouped the data by the reporting years (2017 and 2018), then summarized the data by the total money spent each year. From the chart, we can see that more money we spent in 2018 than 2017

```

Tally the total overall disbursement amounts for each month (distinguishing between say, Oct. 2018 v. Oct. 2019). In which month was spending the highest?

```{r}

payments %>% 
  filter(disbursement_year==2017) %>% 
  group_by(disbursement_month) %>% 
  summarize(tot_money = sum(disbursement_amount))

payments %>% 
  filter(disbursement_year==2018) %>% 
  group_by(disbursement_month) %>% 
  summarize(tot_money = sum(disbursement_amount))
  

# I could not find anything in the dataset for the year 2019, so I used the years 2017 and 2018

# I ran two different programs in order to distinguish between years. From the two tables, we can conclude that October 2018 was the highest spending month.


```

Which recipient (recipient_name) was paid the most in the 2020 election cycle?

```{r}

payments %>% 
  group_by(recipient_name) %>% 
  summarize(tot_money = sum(disbursement_amount)) %>%
  arrange(desc(tot_money))

# In this code I grouped the data by recipient name, had the data coalesced, and finally had the data arranged to easily show the most paied recipient. As a result, we can see that American Made Media Consultants LLC is the highest recipient.

```

Look at what this recipient was paid for. What service or good (see the "disbursement_description") were they paid the most for?

```{r}

payments %>% 
  filter(recipient_name=="AMERICAN MADE MEDIA CONSULTANTS, LLC") %>% 
  group_by(disbursement_description) %>% 
  summarize(tot_money = sum(disbursement_amount)) %>%
  arrange(desc(tot_money))
  
# I filtered the dataset for "AMMC, LLC", then grouped all the data entries by disbursmeent descriptions and had the data summarized. AMMC LLC was paid most for "Placed Media."

```

What's the most common category overall (see "disbursement_description") the campaign spent money on?

```{r}

payments %>% 
  group_by(disbursement_description) %>% 
  summarize(tot_money = sum(disbursement_amount)) %>%
  arrange(desc(tot_money))

#A problem has arisen with the titles- since titles that share a category don't share an exact name, R cannot put them together. Taking a glance at the data, it appears that, by frequency of payments Amex payments for travel are the most common. By total money spent, then the highest category would be digital consulting/online advertising.

```

Who were the 5 biggest recipients of campaign spending on "LEGAL CONSULTING"..?

```{r}

payments %>% 
  filter(disbursement_description=="LEGAL CONSULTING") %>% 
  group_by(recipient_name) %>%
  arrange(recipient_name) %>% 
  summarize(tot_money = sum(disbursement_amount)) %>%
  arrange(desc(tot_money))

# I filtered the data for legal consulting, grouped and summarized it by recipient names, and then arranged the data. We can say the biggest 5 recipients were Jones Day, Larocca, Kasowitz, Alan S. Futerfas, and Mcdermott.

```


Who are the top 5 individuals (vs. organization) were paid by the campaign committee?

```{r}

payments %>% 
  filter(entity_type=="IND") %>% 
  group_by(recipient_name) %>%
  arrange(recipient_name) %>% 
  summarize(tot_money = sum(disbursement_amount)) %>%
  arrange(desc(tot_money))

# I used the code from Question 6, but filtered for individuals instead of for legal consulting. Using this function, I found the top 5 individuals to be Michael Glassner, Sean Dollman, Darren Centinello, Christopher Carr, and John Pence



```


For the indvidual who was paid the most, which payment category (disbursement_description) was the top one for him or her? And how much money were they paid in that category?

```{r}







```
